Gradient Descent for Numerical MAX-CSP

### 2022-05-01 proj/paper/paper.tex
save paper

% TODO

@@

\section{Background}

@@

Constraint satisfaction problems (CSPs) encompasses many different problems. Formally a 
constraint satisfaction problem (CSP) is defined as a triple $\langle X, D, C \rangle$, where 
% XXX this is from wikipedia https://en.wikipedia.org/wiki/Constraint_satisfaction_problem
\begin{align*}
	X &= \{X_1, ..., X_n\} \text{ the set of variables}\\
	D &= \{D_1, ..., D_n\} \text{ the domains of each respective variable}\\
	C &= \{C_1, ..., C_m\} \text{ the set of constraints}\\
\end{align*}
A classic example of a CSP, is the 3-SAT problem. In this case
the boolean variables that appear in at least one clause would be $X$. The domain for 
each variable would be $D_i = \{true, false\}$ and the constraints would all be 
three variable disjunctive clauses. The 3-SAT problem of course known to be NP-Complete.
In this paper we focus on a subset of CSPs called Numerical MAX-CSP. 
In typical a CSP all constraints must be satisfied, however in MAX-CSP the goal is simply to 
satisfy as may constraints as possible, additionally the domains are the variables 
are numerical (e.g. the real numbers).

While this problem has certainly been considered before, we were only able to find one previous 
work that directly addresses this issue. 
%TODO Add citation here

In this paper the authors propose a solution to numerical MAX-CSP. In particular where 
$D_i = \mathds{R}$ and each constraint $C_j = a_j^Tx \leq b_j, a_j \in \mathds{R}^m, b_j \in \mathds{R}$. 
That is, given a set of linear inequalities, satisfy as many as possible. In the paper the authors 
propose a specific algorithm for this problem based on branch and bound techniques. Unfortunately, 
this algorithm is worst case exponential time. In fact, this problem is easy to formulate as a mixed integer 
program,

@@

Where $\epsilon$ is some large number. The general problem of mixed integer linear programming is 
of course NP-Hard in general (by reduction to 0-1 integer programming which is NP-Complete). Hence 
this numerical MAX-CSP is likely not to admit an efficient solution.

@@

Full text search engines are used in a plethora of applications. These search engines (such as Apache Lucene)
are built for efficient retrieval of top-k documents based on a TF/IDF based scoring metric which is dot 
product between two sparse vectors $q^Td = score$[[1]][1][[2]][2][[3]][3]. While this default scoring gives decent results out of the box, 
it is frequently augmented by re-weighting the query $q$, with some weight vector $w$
changing to scoring to be $(q\odot w)^Td = score$. This allows for boosting of certain terms or fields
to improve the quality of search results while not changing the underlying search algorithm.

The problem we will address in this project is finding a good weight vector $w$. In particular our
problem setting is as follows. We are giving a set of query vectors $Q = \{q_1, ... q_n\}$. For each 
of these query vectors $q_i$ we are given a set of $k$ retrieved document vectors with labels $R_i = \{(d_{i,1}, l_{i,1}), ..., (d_{i, k}, l_{i,k})\}$, 
$l \in \{relevant, irrelevant\}$.
We wish to find a weight vector $w$ that minimizes the number of irrelevant documents that are examined 
before the relevant result is found. This problem can be formulated as a mixed integer linear program (MILP)
of the following form
\begin{align*}
\min_{w,z}\quad &1^Tz\\
s.t. \quad Aw - \epsilon z &\leq \gamma\\
		w &\geq 0\\
		z &\in \{0,1\}\\
\end{align*}

@@

As metioned above we can formulate our problem as a mixed integer linear program 
with binary constraints. In addition, to the mixed integer formulation, we also create an LP
relaxation of the problem as an approximation of the MILP, namely,

\begin{align*}
\min_{w,z}\quad &1^Tz\\
s.t. \quad Aw - \epsilon z &\leq \gamma\\
		w &\geq 0\\
		0 \leq z &\leq 1\\
\end{align*}

The rationale behind this is that the LP relaxation can be solved efficiently and provide a 
deterministic approximation of the MILP. Although the solution could be arbitrarily bad,
we find that the LP typically finds a reasonable solution in much less time 
relative to the MILP.

@@

$$
f(x) = \text{HardSigmoid}(Aw - \gamma)
$$
Where $ \gamma = -1$ and, 
$$
x\in R, \quad \text{HardSigmoid}(x) = \begin{cases} 0 \text{ if } x \leq 0\\ 1 \text{ if } x \geq 1\\ x \text{ otherwise}\end{cases}
$$

@@

>>>>

% TODO clean up

@@

\section{Preliminaries}

In this section we will give an overview of constraint satisfaction problems,
and the specific flavor that we are attempting to tackle in this paper. We then 
give a brief overview gradient descent which is the basis of our proposed solution.

@@

Constraint satisfaction problems (CSPs) encompasses many different problems. Formally a 
constraint satisfaction problem (CSP) is defined as a triple $\langle X, D, C \rangle$ \cite{wiki_CSP}, where 
\begin{align*}
	X &= \{X_1, ..., X_n\} \text{ the set of variables}\\
	D &= \{D_1, ..., D_n\} \text{ the domains of each respective variable}\\
	C &= \{C_1, ..., C_m\} \text{ the set of constraints}\\
\end{align*}
A classic example of a CSP, is the 3-SAT problem. In this case
the boolean variables that appear in at least one clause would be $X$. The domain for 
each variable would be $D_i = \{true, false\}$ and the constraints would all be 
three variable disjunctive clauses. 
In this paper we focus on a subset of CSPs called Numerical MAX-CSP. 
In typical a CSP all constraints must be satisfied, however in MAX-CSP the goal is simply to 
satisfy as may constraints as possible, additionally the domains are the variables 
are numerical (e.g. the real numbers).

While this problem has certainly been considered before, we were only able to find one previous 
work that directly addresses this issue \cite{num_max_csp_paper}.
In this paper the authors propose a exact solution to numerical MAX-CSP. In particular where 
$D_i = \mathds{R}$ and each constraint $C_j = a_j^Tx \leq b_j, a_j \in \mathds{R}^m, b_j \in \mathds{R}$. 
That is, given a set of linear inequalities, satisfy as many as possible. In the paper the authors 
propose a specific algorithm for this problem based on branch and bound techniques. While the algorithm 
does produce optimal solutions, the algorithm is worst case exponential time.
In fact, this problem is readily to formulated as a mixed integer linear program (MILP),

@@

Where $\epsilon$ is some large number. The general problem of mixed integer linear programming is 
NP-Hard in general (by reduction to 0-1 integer programming which is NP-Complete \cite{karp_1972}). Hence 
numerical MAX-CSP is likely not to admit an efficient solution.

@@

Full text search engines are used in a plethora of applications. These search engines (such as Apache Lucene)
are built for efficient retrieval of top-k documents based on a TF/IDF based scoring metric which is dot 
product between two sparse vectors $q^Td = score$\cite{lucene_history} \cite{WAND_paper} \cite{block_max_WAND_paper}.
While this default scoring gives decent results out of the box, 
it is frequently augmented by re-weighting the query $q$, with some weight vector $w$
changing to scoring to be $(q\odot w)^Td = score$. This allows for boosting of certain terms or fields
to improve the quality of search results while not changing the underlying search algorithm.

The problem we will address in this project is finding a good weight vector $w$. In particular our
problem setting is as follows. We are giving a set of query vectors $Q = \{q_1, ... q_n\}$. For each 
of these query vectors $q_i$ we are given a set of $k$ retrieved document vectors with labels $R_i = \{(d_{i,1}, l_{i,1}), ..., (d_{i, k}, l_{i,k})\}$, 
$l \in \{relevant, irrelevant\}$.
We wish to find a weight vector $w$ that minimizes the number of irrelevant documents that are examined 
before the relevant result is found. This problem can be formulated as a mixed integer linear program (MILP)
as above but with fixed $\gamma < 0$ instead of $b$, in particular,
\begin{align*}
\min_{w,z}\quad &1^Tz\\
s.t. \quad Aw - \epsilon z &\leq \gamma\\
		w &\geq 0\\
		z &\in \{0,1\}\\
\end{align*}

@@

As metioned above we can formulate our problem as a mixed integer linear program 
with binary constraints. In addition, to the mixed integer formulation, we also create an LP
relaxation of the problem as an approximation of the MILP.
The rationale behind this is that the LP relaxation can be solved efficiently and provide a 
deterministic approximation of the MILP. Although the solution could be arbitrarily bad,
we find that the LP typically finds a reasonable solution in much less time 
relative to the MILP.

@@

$$
f(w) = \text{HardTanh}(Aw - \gamma)
$$
Where $ \gamma = -1$ and, 
$$
x\in R, \quad \text{HardTanh}(x) = \begin{cases} -1 \text{ if } x \leq -1\\ 1 \text{ if } x \geq 1\\ x \text{ otherwise}\end{cases}
$$

@@

\section{Discussion}
\section{Future Work}
